Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-06-02 manuscripts/manuscript.Rmd
tweaks to manuscript; makefile does not restore cache by default

```{r npv_table, dependson="print_npv", results='asis'} 
pandoc.table(print_npv)
```

@@

$$\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^\infty \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}$$

@@

$$\Pi_0(N_t,h_t) = p h_t -  c_0 E_t \label{eq:annual_netrevenue}$$

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t=qE_tN_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. [eq:annual_netrevenue], $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$.

Taken together this objective function ([eq:objective]) and the state
equation ([eq:state_equation]) define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $h_0^{\ast}$.

<!--NOTE presumably you used a finite terminal time in which case
we need to revise to reflect how you handled that.-->

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. [eq:objective].

Taken together with the state equation (Eqn. [eq:state_equation]), each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy $h_i^*$.

<!-- NOTE Will need to do the runs and generate the figures for that one  -->
In addition in the Supplementary Material, we compare our results with a
more conventional fisheries economic formulation in which additional
costs are applied to the control variables themselves, as opposed to
adjustments to the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$. Additional costs of this
form tend to have a smoothing effect on optimal quotas, but also change
the long-term average stock size or quota size that is optimal.

@@

Each of the policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal net
present value of fishery, broadly following the example of Bovenberg,
Goulder and Gurnery (2005) and Bovenberg, Goulder and Jacobsen (2008).
<!--NOTE Jake to confirm these are still relevant and needed. We are using a less sophisticated normalization than was once discussed and may or may not need the citations now. -->

>>>>

```{r npv_table, dependson="print_npv", results='asis'} 
# verify that we have fine enough c_2 sampling to hit about 75% NPV0 on each penalty form
# pandoc.table(print_npv)
```

@@

$$
\begin{equation}
\max_{{\bf h}}\mathbf{ E} ( NPV_{0} )=\max_{{\bf h}} \sum_0^\infty \mathbf{E} \left( \displaystyle \frac{\Pi_0(N_t,h_t)}{(1+\delta)^{t+1}} \right) \label{eq:objective}
\end{equation}
$$

@@

$$\begin{equation}
\Pi_0(N_t,h_t) = p h_t -  c_0 E_t \label{eq:annual_netrevenue}
\end{equation}
$$

where $E_t$ represents fishing effort. We assume catch is proportional
to stock size and effort expended fishing that year, $h_t = q E_t N_t$ and
constant $q>0$ is the catchability coefficient. In
Eqn. \eqref{eq:annual_netrevenue}, $p$ is the price per unit harvest and
$c_0$ the cost of fishing and, to simplify the presentation of results,
we assume that these are constants with $p>0$ and $c_0>0$. (Examples 
shown here use $p$= `r p` and $c_0$ = `r_c0`.) 
<!-- Mention R code provided to test other values? Do we need to test other values?
(i.e. the c0 = 0 case?) Word about units?-->

Taken together this objective function \eqref{eq:objective} and the state
equation \eqref{eq:state_equation} define a stochastic dynamic programming
problem that we solve using backwards recursion via Bellman's equation.
We denote the resulting state-dependent, optimal control as $h_0^{\ast}$.

We solve this problem on a finite time horizon of $T =$ `r OptTime` using
value iteration [@Mangel1988; @Clark2000].

@@

where indicator function $\mathbf{I}(h_t, h_{t-1})$ takes the value 1 if
$h_t=h_{t-1}$ and zero otherwise. For each $\Pi_i$, we can then define a
new objective function $NPV_i$ similar to that in Eqn. \eqref{eq:objective}.

Taken together with the state equation Eqn. \eqref{eq:state_equation}, each
of these new objective functions defines a different stochastic dynamic
programming problem. Again, we solve them numerically using backwards
recursion. To include costs of policy adjustment, we expand the state
space to include both the current stock size and the management action
taken on the previous time step $(N_t,h_{t-1})$. For each new objective
function $\max NPV_i$, we denote the corresponding optimal control
policy $h_i^*$.

In addition in the Supplementary Material, we compare our results with 
more conventional fisheries economic formulation in which additional
costs are applied to the control variables themselves, as opposed to
adjustments to the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$. Additional costs of this form tend to have a smoothing effect on optimal quotas, but also change
the long-term average stock size or quota size that is optimal.

@@

<!--NOTE c_i notation is confusing because it does not usually refer to c_0 or c_4 -->
Each of the policy adjustment cost function is characterized in terms of
a cost coefficient $c_i$. However, $c_i$ takes different units for each
functional form. Therefore, if seeking to compare the relative effect of
each penalty function on optimal management, it is unclear what
parameter values should be used. To address this issue, we calibrate
choices of $c_i$ so that each has a comparable impact on the optimal net
present value of fishery, broadly following the example of Bovenberg,
Goulder and Gurnery (2005) and Bovenberg, Goulder and Jacobsen (2008).
<!--NOTE Jake to confirm these are still relevant and needed. We are using a less sophisticated normalization than was once discussed and may or may not need the citations now. -->
